---
title: "GPS STPE tracks 2019, 2021 - summary info"
date: "08/12/2021"
output: 
  html_document:
    theme: default
    highlight: tango
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
```

## load packages

```{r packages, message=FALSE}
library(tidyverse)
library(lubridate)
library(ggplot2)
library(viridis)
```

```{r, echo=FALSE}
source("R_scripts/functions/set_ggplot_custom_theme.R")
```


## 1. load and check data
```{r}
gps <- readRDS("data_clean/stpe_gps_data_with_tripID.rds")

str(gps)

## Variable Descriptions #############################################

#deployID: band number - tagID - year
#location: site bird tagged 
#tagID: gps tag code
#metalBand: band number
#species: FTSP or LESP
#fixFrequency: number of hours between GPS fixes
#ts: timestamp of the gps fix or attempted fix
#date: date of gps fix or attempted fix
#num_satelites: number of satellites used to obtain a fix
#lat: latitude of gps fix
#lon: longitude of gps fix
#batery: battery level at time of gps fix or attempted fix
#island_lat: tag deployment latitude
#island_lon: tag deployment longitude
#tagAttachment: method of tag attachment, TT = tesa tape, SLT = sutures + topical lidocaine, SLI = sutures + injected lidocaine
#deployYear: year tag deployed
#deployTime: timetstamp of when tag was deployed
#retrievalTime: timestamp of when tag was retrieved (retrieval times were not recorded so set to 6am on date of tag retrieval)
#deployNestStage: nest status (chick or egg) at tag deployment
#islandDist: distance in km of gps fix from deployment location
#tripCat: identifies what part of the trip the fix or attempted fix belongs to (tripStart, tripEnd, tripMiddle)
#completeTrack: a track is considered complete if the last successful fix of a deployment is < 30 km * fixFrequency from the deploy location
#tripTime: the time from the first trip fix to last trip fix in days (only calculated for complete trips)
#firstDetect: timestamp of the last fix of the deployment
#lastDetect: timestamp of the first fix of the deployment
#tripID: used for birds with more than one trip to differentiate different trips recorded during the same deployment

#####################################################################
```

## 2. summary info

### number of birds tracked
```{r}
nrow(gps %>% select(metalBand) %>% distinct()) #79
```

### total number of trips recorded
```{r}
nrow(gps %>% select(deployID, tripID) %>% distinct()) #111
```

### num of complete/incomplete trips by island, species and deployNestStage
```{r, echo=FALSE, message=FALSE}
knitr::kable(gps %>% 
               filter(!is.na(lat)) %>%
               select(deployID, location, species, tripID, deployNestStage, completeTrack) %>%
               distinct() %>%
               group_by(location, species, deployNestStage, completeTrack) %>%
               mutate(completeTrack = ifelse(completeTrack == 0, "no", "yes")) %>%
               summarise(n = n()) %>%
               arrange(desc(completeTrack)))
```

### gps fix success across deployments/trips (complete tracks only)
```{r, echo=FALSE, message=FALSE}
knitr::kable(gps %>% 
               filter(completeTrack == 1) %>%
               mutate(fix = ifelse(is.na(lat), "unable_to_get_fix", "fix_obtained")) %>%
               group_by(deployID, location, species, tripID, fix) %>%
               summarise(n = n()) %>%
               pivot_wider(names_from = fix, values_from = n) %>%
               mutate(unable_to_get_fix = replace_na(unable_to_get_fix, 0),
                      proportion_fixes_successful = round(fix_obtained/(fix_obtained+unable_to_get_fix), digits = 2)) %>%
               as.data.frame())
```


### Settle time
#### approximate time between tag deployment/attachement and a bird's departure from the colony (departure time estimated based on first successful fix of a deployment)
```{r, echo = FALSE, fig.width=10, fig.height=4}
ttd <- gps %>% select(deployID, ts, deployTime, deployNestStage) %>%
  group_by(deployID) %>%
  arrange(ts) %>%
  slice_head() %>%
  mutate(timetoDepart = as.numeric(round(difftime(ts, deployTime, units = "hours"), digits = 2)))

brx <- seq(0, max(ttd$timetoDepart), by = 2)

ggplot() +
  geom_histogram(data = ttd, aes(timetoDepart, fill = deployNestStage),
                 color = "black", breaks = brx) +
  scale_x_continuous(breaks = round(seq(0, max(ttd$timetoDepart), by = 10), 0)) +
  scale_fill_viridis_d(begin = 0.1, end = 0.7) +
  xlab("time between tagging and departure (hours)")

```

### Colony departure and arrival times 
#### times estimated based on first and last fix of each trip
```{r, echo=FALSE, fig.width=10, fig.height=4}
tda <- gps %>%
  dplyr::select(ts, tripCat, completeTrack) %>%
  dplyr::filter(tripCat %in% c("tripStart", "tripEnd") & completeTrack == 1) %>%
  mutate(hms = as.POSIXct(format(ts, format = "%H:%M:%S"), format = "%H:%M:%S", tz = "America/Vancouver"))


ggplot(tda) +
  geom_histogram(aes(x = hms, fill = tripCat), binwidth = 60*60,  color = "black") +
  scale_x_datetime(date_breaks = "4 hour",
                   date_labels = "%H:%M:%S",
                   timezone = "America/Vancouver",
                   expand = c(0, 0)) +
  scale_y_continuous(limits = c(0, 80), expand = c(0,0)) +
  xlab("time") +
  scale_fill_viridis_d(begin = 0.2, end = 1) 
```

### Trip time and maximum distance traveled away from the colony
```{r, echo = FALSE, message=FALSE}

#calculate trip time
tt <- gps %>% 
  filter(!is.na(lat) & !is.na(deployNestStage)) %>%
  dplyr::select(location, deployID, tripID, species, deployNestStage, completeTrack, firstDetect, lastDetect) %>%
  distinct() %>%
  mutate(tripTime = round(difftime(lastDetect, firstDetect, units = "days"), digits = 2),
         deployIDt = paste0(deployID, " (t", tripID,")")) 

# %>%
#   arrange(deployNestStage, location, tripTime) %>%
#   mutate(deployIDt = factor(deployIDt, levels = deployIDt))

## calculate maximum distance traveled from the colony
mdist <- gps %>% 
  filter(!is.na(deployNestStage)) %>%
  group_by(deployID, tripID, location, species, deployNestStage) %>%
  summarise(maxdist = max(islandDist, na.rm = TRUE)) %>%
  mutate(deployIDt = paste0(deployID, " (t", tripID,")")) 

# %>%
#   arrange(deployNestStage, location, maxdist) %>%
#   mutate(deployIDt = factor(deployIDt, levels = deployIDt))

```

#### Trip time 
```{r, echo = FALSE, message=FALSE, fig.width=10, fig.height=4}
ggplot() +
  geom_histogram(data = tt, aes(x = tripTime, fill = location), color = "black", binwidth = 1) +
  facet_grid(.~deployNestStage) +
  scale_y_continuous(limits = c(0, 20), expand = c(0, 0)) +
  xlab("trip time (days)") +
  ylab("# of trips") +
  scale_fill_viridis_d()

```

#### Maximum distance from colony
```{r, echo = FALSE, fig.width=10, fig.height=4}
ggplot() +
  geom_histogram(data = mdist, aes(x = maxdist, fill = location), color = "black", binwidth = 100) +
  facet_grid(.~deployNestStage) +
  scale_y_continuous(limits = c(0, 12), expand = c(0, 0)) +
  xlab("max distance from colony (km)") +
  ylab("# of trips") +
  scale_fill_viridis_d()

```

### Plots of trip time and max distance for individual trips
#### Total trip time for each trip
```{r, echo = FALSE, fig.width=10, fig.height=14}

ggplot() +
  geom_segment(data = tt, aes(x = tripTime, xend = 0, y = deployIDt, yend = deployIDt, color = location), size = 3) +
  facet_grid(deployNestStage~., scales = "free_y") +
  scale_x_continuous(limits = c(0, 9), expand = c(0, 0)) +
  xlab("trip time (days)") +
  ylab("deploy ID (trip #)") +
  scale_color_viridis_d()

```

#### Furthest distance from colony for each trip
```{r, echo = FALSE, fig.width=10, fig.height=14}

ggplot() +
  geom_segment(data = mdist, aes(x = maxdist, xend = 0, y = deployIDt, yend = deployIDt, color = location), size = 3) +
  facet_grid(deployNestStage~., scales = "free_y") +
  scale_x_continuous(limits = c(0, 1500), expand = c(0, 0)) +
  xlab("max distance from colony (km)") +
  ylab("deploy ID (trip #)") +
  scale_color_viridis_d()

```



